\ifdefined\ThesisLanguageIsEnglish
\chapter*{Abstract}
\else
\chapter*{Kurzfassung}
\fi
\label{ch:Abstract}

\begin{abstract}

The increasing penetration of inverter-based resources (IBRs) is fundamentally changing the dynamic behavior of modern power systems, leading to reduced inertia and weaker grid conditions. In this context, grid-forming (GFM) converters have emerged as a promising solution to enhance system stability by providing voltage and frequency support.

This thesis investigates the impact of GFM converters on power system stability under large disturbances using RMS simulations in DIgSILENT PowerFactory. The analysis focuses on key performance indicators, including the rate of change of frequency (RoCoF), frequency nadir, settling time, and voltage recovery.

Additionally, the effect of increasing GFM penetration is assessed using a modified IEEE 39-bus system, as well as its impact on small-signal stability.

The results show that grid-following-dominated systems exhibit reduced stability margins, particularly in weak grids, whereas the integration of GFM converters significantly improves frequency stability and the damping of oscillations. The effectiveness of GFM control is found to be highly dependent on grid conditions. 

Finally, transmission-level analyses based on Power Transfer Distribution Factors (PTDF) and Transfer Capability Analysis (TCA) are performed to evaluate how the integration and location of new generation sources impact power flows and network constraints. These analyses provide complementary insight into how local converter behavior translates into system-wide effects, particularly in terms of congestion and transfer limits.

Overall, this work provides insights into the role of grid-forming converters in future low-inertia power systems and supports their integration through improved system-level assessment.

\end{abstract}
